Attack Graph Generation for Micro-service Architecture

### 2018-09-20 paper_own_docker/chapters/evaluation.tex
TODOS in the method

Modern microservice architectures use an abundance of different technologies,  number of containers, various connectivity and number of vulnerabilities. Therefore it is of immense importance to show that an attack graph generator works well in such heterogeneous scenarios. In order to do this, we tested our system on some real and slightly modified Github examples as described in Table \ref{table_technologies}. Our intention was to find and test examples that are publicly available for possible future comparison characterized by different system properties(topologies, technologies, vulnerabilities) and coming from different usage domains. We also had to take into account that an overwhelming majority of the examples publicly available are small with only one or a few containers, which made this search challenging. The resulting examples are as follows: NetflixOSS \cite{netflixoss}, Atsea Sample Shop App \cite{atsea} and JavaEE demo \cite{java}. NetflixOSS is a microservice system provided by Netflix that is composed of 10 containers and uses Spring Cloud, Netflix Ribbon, Netflix Eureka etc. Atsea Sample Shop App is an e-commerce sample web application composed of 4 containers and that uses Spring Boot, React, NGINX and PostgreSQL. JavaEE demo is a sample application for browsing movies that is composed of only two containers and uses JavaEE, React and Tomcat EE. We ran the attack graph generator and verified the resulting attack graphs of the small examples manually based on domain knowledge and under the assumption that the output from Clair \cite{clair}, NVD attack vectors \cite{booth2013national} and the pre- and postconditions from the work of Aksu et al. \cite{aksu2018automated} are correct. After running the attack graph generator, the attack graphs for the Atsea Sample Shop app and the JavaEE demo are small as expected with few nodes and edges. The structure of the resulting Netflix attack graph had a nearly linear structure in which each node is connected to a small number of other nodes that form a chain of attacks. This linearity is because each container is connected to a few other containers to reduce unnecessary communication and increase encapsulation. Therefore based on this connectivity an attacker needs to perform multiple intermediate steps in order to reach the target container. All of the examples terminated, there are no directed edges from containers with higher privileges to lower privileges, no duplication of nodes and no reflexive edges, which is in line with the previously mentioned monotonicity property. Additionally, we noticed that the running time of our system for each of these examples was short, and additional scalability tests are needed. The Phpmailer \cite{phpmailer} and Samba \cite{samba} system is an artificial example that we use and extend in the following subsection to perform these scalability tests.

\begin{table*}
	\begin{center}
		\begin{tabular}{ cccccc } 
			\hline
			Statistics & example\_20 & example\_50 & example\_100 & example\_500 & example\_1000 \\ \hline

>>>>

Modern microservice architectures use an abundance of different technologies,  number of containers, various connectivity and number of vulnerabilities. Therefore it is of immense importance to show that an attack graph generator works well in such heterogeneous scenarios. In order to do this, we tested our system on some real and slightly modified Github examples as described in Table \ref{table_technologies}. Our intention was to find and test examples that are publicly available for possible future comparison characterized by different system properties(topologies, technologies, vulnerabilities) and coming from different usage domains. We also had to take into account that an overwhelming majority of the examples publicly available are small with only one or a few containers, which made this search challenging. The resulting examples are as follows: NetflixOSS \cite{netflixoss}, Atsea Sample Shop App \cite{atsea} and JavaEE demo \cite{java}. NetflixOSS is a microservice system provided by Netflix that is composed of 10 containers and uses Spring Cloud, Netflix Ribbon, Netflix Eureka etc. Atsea Sample Shop App is an e-commerce sample web application composed of 4 containers and that uses Spring Boot, React, NGINX and PostgreSQL. JavaEE demo is a sample application for browsing movies that is composed of only two containers and uses JavaEE, React and Tomcat EE. We ran the attack graph generator and verified the resulting attack graphs of the small examples manually based on domain knowledge and under the assumption that the output from Clair \cite{clair}, NVD attack vectors \cite{booth2013national} and the pre- and postconditions from the work of Aksu et al. \cite{aksu2018automated} are correct. After running the attack graph generator, the attack graphs for the Atsea Sample Shop app and the JavaEE demo are small as expected with few nodes and edges. The structure of the resulting Netflix attack graph had a nearly linear structure in which each node is connected to a small number of other nodes that form a chain of attacks. This linearity is because each container is connected to a few other containers to reduce unnecessary communication and increase encapsulation. Therefore based on this connectivity an attacker needs to perform multiple intermediate steps in order to reach the target container. All of the examples terminated, there are no directed edges from containers with higher privileges to lower privileges, no duplication of nodes and no reflexive edges, which is in line with the previously mentioned monotonicity property. Additionally, we noticed that the running time of our system for each of these examples was short, and additional scalability tests are needed. The Phpmailer \cite{phpmailer} and Samba \cite{samba} system is an artificial example that we use and extend in the following subsection to perform these scalability tests.
\begin{table*}[t]
	\begin{center}
		\begin{tabular}{ p{20mm}p{35mm}p{35mm}p{10mm}cp{35mm} } 
			\hline
			Name & Description & Technology stack & No. Containers & No. vuln. & Github link \\\hline

			Netflix OSS & Combination of containers provided from Netflix. & Spring Cloud, Netflix Ribbon, Spring Cloud Netflix, Netflix's Eureka & 10 & 4111 & \url{https://github.com/Oreste-Luci/netflix-oss-example} \\

			Atsea Sample Shop App & An example online store application. & Spring Boot, React, NGINX, PostgreSQL & 4 & 120 & \url{https://github.com/dockersamples/atsea-sample-shop-app} \\

			JavaEE demo & An application for browsing movies along with other related functions. & Java EE application, React, Tomcat EE & 2 & 149 & \url{https://github.com/dockersamples/javaee-demo} \\

			PHPMailer and Samba & An artificial example created from two separate containers. We use an augmented version for the scalability tests. & PHPMailer(email creation and transfer class for PHP), Samba(SMB/CIFS networking protocol) & 2 & 548 &  \url{https://github.com/opsxcq/exploit-CVE-2016-10033}
			\url{https://github.com/opsxcq/exploit-CVE-2017-7494} \\

			\hline
		\end{tabular}
	\end{center}

	\caption{Microservice architecture examples analyzed by the attack graph generator}
	\label{table_technologies}

\end{table*}
\begin{table*}
	\begin{center}
		\begin{tabular}{ cccccc } 
			\hline
			Statistics & example\_20 & example\_50 & example\_100 & example\_500 & example\_1000 \\ \hline

### 2018-09-20 paper_own_docker/chapters/method.tex
remove TODO
TODOS in the method

\subsection{From Network Nodes to Microservices}
\label{chap:mapping}
% abit reptative 
In our work, we adapt already existing attack graph generation methods from computer networks to the microservices ecosystem. In order to do this, we identify the different components and find a compatible replacement that can be used in a microservice architecture. In this subsection, we start first by shortly introducing a famous microservice framework (Docker) and some of its terminology. We then modify the attack graph concepts mentioned in Subsection \ref{chap:attack_graphs}: nodes, edges, privilege levels, pre- and postconditions to match our use-case. We illustrate the whole idea  by demonstrating a small example.

@@

Privileges play a central role in the generation if attack graphs. Traditionally, the privileges are modeled as a hierarchy that varies in the access type (\textit{User, Admin}), and the access scope (Virtual machine VOS, host machine OS). The exhaustive list of privileges, that are used in this paper, are: \textit{None, VOS(User), VOS(Admin), OS(User) and OS(Admin)}. VOS means that the privilege is exclusive to a virtual machine, while not affecting the host machine. However in our case, unlike hosts connected in a computer network, these privileges are adapted to images and not virtual machines. On the other side, the keyword OS means that the  host machine can be controlled by a user who has this privilege. Since VOS are isolated from host machines and their exploitation does not imply the exploitation of the host machine, they are in the lower level of hierarchy \cite{aksu2018automated}. None means that no privilege is obtained, User means only a subset of user level privileges are available, while Admin grants control over the whole system.

As mentioned earlier, \textit{nodes} and \textit{edges} are the basic building blocks of an attack graph. \textit{Nodes} are represented as a combination of docker images and their respective compromise levels (presented as privileges granted to the attacker), while directed edges represent attacks from start nodes (images with their current privileges) to end nodes (exploitable adjacent images with gained privileges). In addition, vulnerabilities that could be exploited in the end nodes serve as edge descriptors.

In order for an attacker to exploit a given vulnerability, certain preconditions in the start node have to be met. Once an attacker meets these preconditions and exploits the vulnerability, he gains the privilege of the end node as a postcondition and a directed edge is added between them. Both the pre- and postconditions in this work are transformed from pre- and postcondition rules manually selected and evaluated by experts in existing work \cite{aksu2018automated}. The pre- and postcondition rules use the fields defined by NVD, as well as an occurrence of specific keywords from the CVEs descriptions \cite{booth2013national}.

@@

\textbf{TODO shorten all the captions, they should be one line, and add all the description to the text. this goes to all the figures and table..}\\
\begin{figure*}[!h]
	\centering
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=.5\linewidth]{./images/Topology_graph}
		\caption{}
		\label{TopologyGraph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=.5\linewidth]{./images/Attack_graph}
		\caption{}
		\label{AttackGraph}
	\end{subfigure}

@@

In order to show how the attack graph generation works in practice, we present a small example. The example is taken from the Netflix OSS Github repository. Netflix OSS example is a Spring Cloud-based microservices architecture that enables Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul) and Client Side Load Balancing (Ribbon) \cite{netflixoss, springcloudnetflix}. Displayed in Figure \ref{TopologyGraph} is a subset of the example topology where each node denotes a container and each edge denotes a connection between two containers. The topology consists of "Outside" node, "Docker daemon" node, Zuul, Eureka and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resiliency, security, and more \cite{netflixzuul}, while Eureka is a REST (Representational State Transfer) based service that is primarily used in the AWS cloud for locating services for the purpose of load balancing and failover of middle-tier servers \cite{netflixeureka}. In Figure \ref{AttackGraph} we can see a part of the resulting attack graph, where a node corresponds to a pair of container plus its privilege, while an edge represents an atomic attack. Parts of both graphs have been intentionally omitted to reduce complexity. An example path that an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gain USER privilege.  Then with this USER privilege, it can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain ADMIN privilege. Once the ADMIN privilege has been obtained on Zuul, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain ADMIN privilege. It is important to note that this is not the only path that the attacker can take in order to have ADMIN privileges on Eureka. Another path would be to exploit the CVE-2018-1124 vulnerability via creating entries in procfs by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only USER privilege on Zuul to gain directly ADMIN privileges of the Eureka container. Our attack graph generator shows both paths since it is of an interest to see every possible route in which a container can be compromised.

\subsection{Attack Graph Generation for Docker Networks}
\label{chap:technical}
 Figure \ref{AttackGraphSystem} gives an overview of the attack graph generator system. In the figure, the rectangles denote the main components of the system, while the arrows describe the flow of the system and the files are the intermediate products. Our attack graph generator is composed of three main components: Topology Parser, Vulnerability Parser and Attack Graph Parser. The Topology Parser reads the underlying topology of the system and converts it into to a format needed for our Attack Graph Parser, the Vulnerability Parser generates the vulnerabilities for each of the images and the Attack Graph Parser generates the attack graph from the topology and vulnerabilities files.

In the following subsections, we first have a look into the system requirements, then describe each of the parsers in more detail and finally examine the characteristics of the Breath-first Search graph transversal algorithm.

\begin{figure}
	\includegraphics[scale=0.9]{./images/AttackGraphSystem}
	\caption{Overview of the Attack Graph Generator System}
	\label{AttackGraphSystem}
\end{figure}

Our system is developed for Docker 17.12.1-ce and Docker Compose 1.19.0 \cite{merkel2014docker}. Docker Compose is a tool for defining and running multi-container Docker applications \cite{dockercompose}. It provides a static configuration file that specifies the system containers, networks, ports etc. The code is written in Python 3.6, and we use Clair \cite{clair} and Clairctl \cite{clairctl} for vulnerabilities scanning.

We developed this system to be used with a specific version of Docker and Docker-Compose. This modular structure could, however, be easily extended to other versions of Docker-Compose, vulnerability scanners and microservice architectures by replacing the Vulnerability and Topology Parsers with custom ones and making sure that their outputs conform to the input of the Attack Graph Parser.

@@

In order to generate an attack graph of a given system, we require an arrangement of its components and connections described by a system topology. The topology of Docker containers can be described at either runtime or statically by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose as our main tool. Docker Compose provides a docker-compose.yml file which is used for extraction of the topology of the system. However different versions of docker-compose.yml, use different syntax. For example, older versions use the deprecated keyword "link", while newer ones use exclusively "networks", to denote a connection between two containers. In this work, we use the keyword "networks" as an indicator that a connection between two containers exists.

However, in the majority of cases, in order for an application to be useful, it has to communicate with the outside world. This is usually done by using publishing ports. This is the case in both computer networks, as well as in microservice architectures.

Another consideration that we take into account is the privileged access. Some containers require certain privileges over the docker daemon in order to function properly. For example, a user may want to run some hardware (web-cam \cite{webcam}) or some applications that demand higher privilege levels (Libvirt \cite{libvirt}) from Docker. In Docker, this is usually done either by mounting the Docker socket or specifying the keyword "privileged" in the docker-compose.yml file. An attacker with access to these containers has also access to the Docker daemon. Once the attacker has access to the Docker daemon, he has potential access to the whole microservice system, since every container is controlled and hosted by the daemon.

@@

In the preprocessing step, we use Clair to generate the vulnerabilities of a given container. Clair is a vulnerability scanner that inspects a Docker image and generates its vulnerabilities by providing CVE-ID, description and attack vector for each vulnerability \cite{clair}. An attack vector is an entity that describes which conditions and effects are connected to this vulnerability. The fields in the attack vector as described by the National Vulnerability Database(NVD) \cite{booth2013national} are: Access Vector (Local, Adjacent Network and Network), Access Complexity (Low, Medium, High), Authentication (None, Single, Multiple), Confidentiality Impact (None, Partial, Complete), Integrity Impact (None, Partial, Complete) and Availability Impact (None Partial, Complete). Unfortunately, Clair does not provide with a ready to use interface to analyze a docker image. As a result, we use Clairctl \cite{clairctl} in order to analyze a complete docker image.

@@

After the topology file is extracted and the vulnerabilities for each container are generated, we continue with the attack graph generation. Here, we first preprocess the vulnerabilities and convert them into sets of pre- and postconditions. In order to do this, we match the attack vectors acquired earlier from the vulnerability database and keywords of the descriptions of each vulnerability to generate attack rules. When a subset of attack vector fields and description keywords matches a given rule, we use the pre- or postcondition of that rule. If more than one rule matches, we take the one with the highest privilege level for the preconditions and the lowest privilege level for the postconditions. If no rule matches, we take None as a precondition and ADMIN(OS) as a postcondition. This results in a list of container vulnerabilities with their preconditions and postconditions.

@@

After the preprocessing step is done, the vulnerabilities are parsed and their pre- and postconditions are extracted. Together with the topology, they are feed into the Breadth-first Search algorithm (BFS).
Breadth-first Search is a popular search algorithm that traverses a graph by looking first at the neighbors of a given node, before diving deeper into the graph. Pseudocode of our modified Breadth-first Search is given in Algorithm \ref{BFSalgorithm}. The algorithm requires a topology and a dictionary of the exploitable vulnerabilities as an input and the output is made up of nodes and edges that make the attack graph. The algorithm first initializes the nodes, edges, queue and the passed nodes. Afterward, it generates the nodes which are a combination of the image name and the privilege level. Then into a while loop, it iterates through every node, checks its neighbors and adds the edges if the conditions are satisfied. If the neighbor was not passed, then it is added to the queue. The algorithm terminates when the queue is empty. Furthermore, BFS is characterized by the following properties:

\begin{table*}[t]
	\begin{center}
		\begin{tabular}{ p{20mm}p{35mm}p{35mm}p{10mm}cp{35mm} } 
			\hline
			Name & Description & Technology stack & No. Containers & No. vuln. & Github link \\\hline

			Netflix OSS & Combination of containers provided from Netflix. & Spring Cloud, Netflix Ribbon, Spring Cloud Netflix, Netflix's Eureka & 10 & 4111 & \url{https://github.com/Oreste-Luci/netflix-oss-example} \\

			Atsea Sample Shop App & An example online store application. & Spring Boot, React, NGINX, PostgreSQL & 4 & 120 & \url{https://github.com/dockersamples/atsea-sample-shop-app} \\

			JavaEE demo & An application for browsing movies along with other related functions. & Java EE application, React, Tomcat EE & 2 & 149 & \url{https://github.com/dockersamples/javaee-demo} \\

			PHPMailer and Samba & An artificial example created from two separate containers. We use an augmented version for the scalability tests. & PHPMailer(email creation and transfer class for PHP), Samba(SMB/CIFS networking protocol) & 2 & 548 &  \url{https://github.com/opsxcq/exploit-CVE-2016-10033}
			\url{https://github.com/opsxcq/exploit-CVE-2017-7494} \\

			\hline
		\end{tabular}
	\end{center}

	\caption{Microservice architecture examples analyzed by the attack graph generator}
	\label{table_technologies}

\end{table*}

\begin{itemize}
	\item Completeness: Breadth-first Search is complete i.e. if there is a solution, Breadth-first search will find it regardless of the kind of graph.
	\item Termination: This follows from the monotonicity property. Monotonicity is ensured if it is assumed that an attacker will never need to relinquish a state \cite{ingols2006practical, ou2006scalable, ammann2002scalable}. In this implementation, each edge is traversed only once, making sure that monotonicity is preserved.
	\item Time Complexity: is $O(|N| + |E|)$ where $|N|$ is the number of nodes and $|E|$ is the number of edges in the attack graph.
\end{itemize}

>>>>

\subsection{From Network Nodes to Microservices}
\label{chap:mapping}
% abit reptative 
In our work, we adapt already existing attack graph generation methods from computer networks to the microservices ecosystem. In order to do this, we identify the different components and find a compatible replacement that can be used in a microservice architecture. In this subsection, we start first by shortly introducing a famous framework (Docker) and some of its terminology. We then modify the attack graph concepts mentioned in Subsection \ref{chap:attack_graphs}: nodes, edges, privilege levels, pre- and postconditions to match our use-case. We illustrate the whole idea  by demonstrating a small example.

@@

Privileges play a central role in the generation if attack graphs. Traditionally, the privileges are modeled as a hierarchy that varies in the access level (\textit{User, Admin}), and the access scope (virtual machine VOS, host machine OS). The exhaustive list of privileges, that are used in this paper, are: \textit{None, VOS(User), VOS(Admin), OS(User) and OS(Admin)}. VOS means that the privilege is exclusive to a virtual machine, while not affecting the host machine. However in our case, unlike hosts connected in a computer network, these privileges refer to images and not virtual machines. On the other side, the keyword \textit{OS} means that the  host machine can be controlled by a user who has this privilege. Since \textit{VOS} are isolated from host machines and their exploitation does not imply the exploitation of the host machine, they are in the lower level of the hierarchy \cite{aksu2018automated}. \textit{None} means that no privilege is obtained, \textit{User} means only a subset of user level privileges are granted, and \textit{Admin} grants control over the whole system.

As mentioned earlier, the \textit{nodes} and the \textit{edges} are the basic building blocks of an attack graph. A \textit{node} represents  a combination of a docker image and its respective compromise levels (expressed as privileges obtained by the attacker). A directed \textit{edge} between two nodes represents an attack step from one node (a compromised image with a certain privilege gained by the attacker) to another node (adjacent exploitable image with the gained privileges). Each edge is typed with the vulnerability (CVE) that could be exploited in the end node.

In order for attackers to exploit a given vulnerability, they need to have certain \textit{preconditions}, i.e, the minimum privileges needed to exploit \cite{aksu2018automated}. Once an attacker meets these preconditions and exploits the vulnerability, he gains the privilege of the end node as a \textit{postcondition} and a directed edge is added between them. Both the pre- and postconditions in this work are transformed from pre- and postcondition rules manually selected and evaluated by experts in existing work \cite{aksu2018automated}. The pre- and postcondition rules use the fields defined by NVD, as well as an occurrence of specific keywords from the CVEs descriptions \cite{booth2013national}.

@@

\begin{figure*}[!h]
	\centering
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=.5\linewidth]{./images/Topology_graph}
		\caption{}
		\label{TopologyGraph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{\columnwidth}
		\centering
		\includegraphics[width=.5\linewidth]{./images/Attack_graph}
		\caption{}
		\label{AttackGraph}
	\end{subfigure}

@@

In order to show how the attack graph generation works in practice, we present a small example. The example is taken from the Netflix OSS Github repository. Netflix OSS example is a Spring Cloud-based microservices architecture that uses the following microservices: Service Discovery (Eureka), Circuit Breaker (Hystrix), Intelligent Routing (Zuul) and Client Side Load Balancing (Ribbon) \cite{netflixoss, springcloudnetflix}. Displayed in Figure \ref{TopologyGraph} is a subset of the example topology where each node denotes a container and each edge is a connection between two containers, if one calls the other. The topology consists of an "Outside" node, "Docker daemon" node, Zuul, Eureka and other nodes. According to Netflix, Zuul is an edge service that provides dynamic routing, monitoring, resiliency, and security functionalities. Eureka is a REST (Representational State Transfer) based service that is primarily used in the cloud for locating services for the purpose of load balancing and fail-over of middle-tier servers. In Figure \ref{AttackGraph} we can see a part of the corresponding attack graph, where a node is a pair of the image and its privilege, while an edge represents an atomic attack. Parts of both graphs have been intentionally omitted to reduce complexity. An example path that an attacker would take could be to first attack the Zuul container by exploiting the CVE-2016-10249 vulnerability by crafting an image file, which triggers a heap-based buffer overflow\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2016-10249}} and gain USER privilege.  With this USER privilege, an attacker can exploit the CVE-2015-7554 vulnerability on the same container via crafted field data in an extension tag in a TIFF image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2015-7554}} to gain ADMIN privilege. Once the ADMIN privilege has been obtained on Zuul, the attacker can attack the Eureka container by exploiting CVE-2017-7600 via another crafted image\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2017-7600}} and gain ADMIN privilege. It is important to note that this is not the only path that the attacker can take in order to have ADMIN privileges on Eureka. Another path would be to exploit the CVE-2018-1124 vulnerability via creating entries in procfs by starting processes, which could result in crashes or arbitrary code execution\footnote{\url{https://nvd.nist.gov/vuln/detail/CVE-2018-1124}}. This vulnerability can be exploited by having only USER privilege on Zuul to gain directly ADMIN privileges of the Eureka container. Our attack graph generator shows both paths since it is of an interest to see every possible route in which a container can be compromised.

\subsection{Attack Graph Generation for Docker Networks}
\label{chap:technical}

 Figure \ref{AttackGraphSystem} gives an overview of the attack graph generator. In the figure, the rectangles denote the main components of the system, while the arrows describe the flow of the system and the files are the intermediate products. Our attack graph generator is composed of three main components: \textit{Topology Parser, Vulnerability Parser and Attack Graph Generator}. The Topology Parser reads the underlying topology of the system and converts it into to a format needed for our Attack Graph Generator, the Vulnerability Parser scans the vulnerabilities for each of the images and the Attack Graph Parser generates the attack graph from the topology and vulnerabilities files. In the following subsections, we first have a look into the system requirements, then describe each component in more details.

The generator is developed and tested for Docker 17.12.1-ce and Docker Compose 1.19.0 \cite{merkel2014docker}. Docker Compose\footnote{\url{https://docs.docker.com/compose/}} is a tool for defining the orchestration of a multi-container applications. It provides a static configuration file that specifies the system containers, networks, and ports. Clair and ClairCtl \footnote{\url{https://github.com/coreos/clair}}  are used for vulnerabilities scanning. The generator is written in Python 3.6. Although we used  specific versions of the tools, the pipe and filter structure of the generator can be easily extended to other versions of Docker-Compose, vulnerability scanners and microservice architectures by replacing.

\textbf{TODO rename the attack graph parser in the box to Attack graph Generator}
\begin{figure}
	\includegraphics[scale=0.9]{./images/AttackGraphSystem}
	\caption{Overview of the Attack Graph Generator System}
	\label{AttackGraphSystem}
\end{figure}

@@

In order to generate an attack graph of a given system, we require an arrangement of its components and connections described as a system topology. The topology of Docker containers can be described either at  runtime or at design time by using Docker Compose. In our case, since we are doing static attack graph analysis, we use Docker Compose to extract the topology. Docker Compose provides a  file (docker-compose.yml) which is used for describes the orchestration of the services. However different versions of docker-compose.yml, use different syntax. For example, older versions use the deprecated keyword "\textit{link}", while newer ones use exclusively "\textit{networks}", to denote a connection between two images. In this work, we use the keyword "\textit{networks}" as an indicator that a connection between two images exists.

In the majority of cases, for an application to be useful, it communicates with the outside world, i.e, it has end points that can be used by the outer network. In Docker, this is usually done by using publishing ports. This is the case in both computer networks, as well as in microservice architectures.

Another consideration that we take into account is the \textit{privileged access} \footnote{\url{http://obrown.io/2016/02/15/privileged-containers.html}} \cite{libvirt}. Some containers obtain certain privileges that grant them control over the docker daemon in order to function properly. For example, a user may want to run some hardware (e.g., web-cam) or some applications that demand higher privilege levels  from Docker. In Docker, this is usually done either by mounting the Docker socket or specifying the keyword "privileged" in the docker-compose.yml file. An attacker with access to these containers has also access to the Docker daemon. Once the attacker has access to the Docker daemon, he has potential access to the whole microservice system, since every container is controlled and hosted by the daemon.

In the preprocessing step, we use Clair to generate the vulnerabilities of a given image. Clair is a vulnerability scanner that inspects a Docker image and generates its vulnerabilities by providing \textit{CVE-ID}, a description and attack vector for each vulnerability. An attack vector is an entity that describes which conditions and effects are connected to this vulnerability. We collect the fields in the attack vector which are,  as described by the National Vulnerability Database(NVD) \cite{booth2013national}, Access Vector (Local, Adjacent Network and Network), Access Complexity (Low, Medium, High), Authentication (None, Single, Multiple), Confidentiality Impact (None, Partial, Complete), Integrity Impact (None, Partial, Complete) and Availability Impact (None Partial, Complete). Unfortunately, Clair does not provide a command line interface to analyze a docker image. We use another tool, i.e.,  Clairctl  to analyze a complete docker image.

@@

@@

After the topology  is extracted and the vulnerabilities for each container are generated, we continue with the attack graph generation. Here, we first pre-process the vulnerabilities and convert them into sets of pre- and postconditions. In order to do this, we match the attack vectors acquired earlier from the vulnerability database and keywords of the descriptions of each vulnerability to generate attack rules. When a subset of attack vector fields and description keywords matches a given rule, we use the pre- or postcondition of that rule. If more than one rule matches, we take the one with the highest privilege level for the preconditions and the lowest privilege level for the postconditions. If no rule matches, we take None as a precondition and ADMIN(OS) as a postcondition. This results in a list of container vulnerabilities with their preconditions and postconditions. \textbf{TODO give an example of the attack rule and refer to the source of these rules}

@@

After the preprocessing step is done, the vulnerabilities are parsed and their pre- and postconditions are extracted. Together with the topology, they are feed into a Breadth-first Search algorithm (BFS).
Breadth-first Search is a popular search algorithm that traverses a graph by looking first at the neighbors of a given node, before diving deeper into the graph. A pseudo-code of our modified Breadth-first Search is given in Algorithm \ref{BFSalgorithm}. The algorithm requires a topology and a dictionary of the exploitable vulnerabilities as an input and the output is made up of nodes and edges that make the attack graph \textbf{TODO please explain what is cont expl, priv acc and refer to them in the text}. The algorithm first initializes the nodes, edges, queue and the passed nodes. Afterward, it generates the nodes which are a combination of the image name and the privilege level. Then into a while loop, it iterates through every node, checks its neighbors and adds the edges if the conditions are satisfied. If the neighbor was not passed, then it is added to the queue. The algorithm terminates when the queue is empty. Furthermore, BFS is characterized by the following properties.
\textbf{TODO: add line numbers to the algo and refer to those numbers in the text.}

\begin{itemize}
	\item Completeness: Breadth-first Search is complete i.e. if there is a solution, Breadth-first search will find it regardless of the kind of graph.
	\item Termination: This follows from the monotonicity property. Monotonicity is ensured if it is assumed that an attacker will never need to relinquish a state \cite{ingols2006practical, ou2006scalable, ammann2002scalable}. In this implementation, each edge is traversed only once, making sure that monotonicity is preserved.
	\item  Complexity: is $O(|N| + |E|)$ where $|N|$ is the number of nodes and $|E|$ is the number of edges in the attack graph.
\end{itemize}
